The prime of their life - the halflife of acting talent

### 2017-08-29 01_actor_careers/05_report.Rmd
director code added

This got me thinking about the rest of the dataset. Given our **12,258** films dataset isn't limited to director credits in our top 250 (films in the set are contributed by both actor and director credits) we can get a bit of a better idea of the broader gender split of directors. I've captured a lot of the key numbers in the following chart.

```{r}
# Directors

# Top 250
top250 = films %>%
          select("name" = director) %>%
          inner_join(film_folk) %>%
          group_by(gender) %>%
          summarise(
            films = n(),
            directors = n_distinct(name)
          )

>>>>

This got me thinking about the rest of the dataset. Given our **12,258** film dataset isn't limited to director credits in our top 250 (films in the set are contributed by both actor and director credits) we can get a bit of a better idea of the broader gender split of directors. I've captured a lot of the key numbers in the following chart.

```{r}
# Establish a persistent dataset of interest
data = films %>%
        select(title, year, tomatometer, director:actor_6) %>%
        gather("role", "name", director:actor_6) %>%
        inner_join(film_folk) %>%
        select(-first_name) %>%
        mutate(
          role_type = ifelse(role == "director", "Director", "Actor"),
          age_at_production = year - dob_year,
          category = case_when(
            role_type == "Director" ~ "director",
            gender == "male" ~ "male_actor",
            gender == "female" ~ "female_actor",
            TRUE ~ "other"
          )
        )

# Directors

# Data for known individuals
known_summary = 
  data %>%
  group_by(role_type, gender) %>%
  summarise(
    films = n(),
    distinct_people = n_distinct(name)
  )

# Unknown directors
unknown_set = films %>%
          select("name" = director) %>%
          anti_join(film_folk) %>%
          mutate(
            first_name = str_extract(name, "(?i)[a-z]+(?=\\s)"),
            min_year = 1900, max_year = 1980
          ) %>%
          filter(!is.na(first_name))

unknown_genders = 
  unknown_set %>% 
  gender_df(name_col = "first_name", year_col = c("min_year", "max_year")) %>%
  select("first_name" = name, gender)

unknown_summary = 
  unknown_set %>%
  inner_join(unknown_genders) %>%
  mutate(role_type = "Director (Non Top250)") %>%
  group_by(role_type, gender) %>%
  summarise(
    films = n(),
    distinct_people = n_distinct(name)
  )

chart_data = 
  known_summary %>%
  union_all(unknown_summary) %>%
  group_by(role_type) %>%
  mutate(
    films_prop = films / sum(films),
    people_prop = distinct_people / sum(distinct_people)
  ) %>% ungroup()

map(c("Actor", "Director", "Director (Non Top250)"), function(x) {

    chart_data %>%
    filter(role_type == x) %>%
    gather("variable", "value", films_prop, people_prop) %>%
    hchart(hcaes(x = gender, group = variable, y = value)) %>%
    hc_title(text = x)

}) %>%
hw_grid(ncol = 1) %>% 
browsable()
